Modal Logics in the Coq Proof Assistant

### 2014-12-09 Papers/InProgress/ModalLogicCoq/ModalCoq.tex
Add image for natural deduction proof of a lemma

\usepackage{fancybox} % Must be before "fancyvrb"
\usepackage[utf8]{inputenc}
\usepackage{url,amsmath,amssymb,fancyvrb,color}
\usepackage{proof}
\usepackage{xspace}

@@

\section{Convenient Modal Lemmas}

\noindent 
The following lemmas are quite convenient. The first is a
kind of modus ponens applicable to formulas under the scope of a modal
operator. The second allows pushing negations inside modalities.

\begin{Verbatim}[commandchars=\\\{\},fontsize=\verbsize]
\Lemma mp_dia: 
  [mforall p, mforall q, (dia p) m-> (box (p m-> q)) m-> (dia q)].
\Proof. mv. 
intros p q H1 H2. dia_e H1. dia_i w0. box_e H2 H3. apply H3. exact H1. 
\Qed.

\Lemma not_dia_box_not: [mforall p, (m~ (dia p)) m-> (box (m~ p))].
\Proof. mv. 
intro p. intro H. box_i. intro H2. apply H. dia_i w0. exact H2. 
\Qed.
\end{Verbatim}

@@

>>>>

\usepackage{fancybox} % Must be before "fancyvrb"
\usepackage[utf8]{inputenc}
\usepackage{url,amsmath,amssymb,fancyvrb,color,graphicx}
\usepackage{proof}
\usepackage{xspace}

@@

\section{Two Simple Modal Lemmas}

In order to illustrate the tactics described above, 
we show \Coq proofs for two simple but useful modal lemmas. 
The first lemma is a kind of modus ponens with formulas 
under the scope of modal operators.

\begin{Verbatim}[commandchars=\\\{\},fontsize=\verbsize]
\Lemma mp_dia: 
  [mforall p, mforall q, (dia p) m-> (box (p m-> q)) m-> (dia q)].
\Proof. mv. 
intros p q H1 H2. dia_e H1. dia_i w0. box_e H2 H3. apply H3. exact H1. 
\Qed.
\end{Verbatim}

% \[
% \infer[\all_I,\all_I]{w: \all p. \all q.  \pos p \imp (\nec (p \imp q)) \imp \pos q}{
%   \infer[\imp_I^1,\imp_I^2]{w: \pos p \imp (\nec (p \imp q)) \imp (\pos q)}{
%     \infer[\pos_I]{w: \pos q}{
%       \infer[\imp_E]{w_0: q}{
%         \infer[\pos_E]{w_0: p}{ 
%           \infer[1]{w: \pos p}{}
%         } 
%         & 
%         \infer[\nec_E]{w_0: p \imp q}{
%           \infer[2]{w: \nec (p \imp q)}{}
%         }
%       }
%     }
%   }
% }
% \]

% \newcommand{\down}[1]{\begin{array}{c}\phantom{p} \\ #1 \end{array} }
% \newcommand{\up}[1]{\begin{array}{c}#1 \\ \phantom{p} \end{array} }

% \[
% \infer[\all_I,\all_I]{\all p. \all q.  \pos p \imp (\nec (p \imp q)) \imp \pos q}{
%   \infer[\imp_I^1,\imp_I^2]{\pos p \imp (\nec (p \imp q)) \imp (\pos q)}{
%     \infer[\pos_I]{\pos q}{
%       \infer[\imp_E]{\up{q}}{
%         \infer[\pos_E]{\down{p} }{ 
%           \infer[1]{\pos p}{}
%         } 
%         & 
%         \infer[\nec_E]{ \down{p \imp q} }{
%           \infer[2]{\nec (p \imp q)}{}
%         }
%       }
%     }
%   }
% }
% \]

\begin{figure}
\centering
\label{fig:mp_dia}
\includegraphics[width=0.5\textwidth]{ND-(mp_dia).png}
\caption{Natural deduction proof of \texttt{mp\_dia}}
\end{figure}

\noindent
The second lemma allows negations to be pushed inside modalities.

\begin{Verbatim}[commandchars=\\\{\},fontsize=\verbsize]
\Lemma not_dia_box_not: [mforall p, (m~ (dia p)) m-> (box (m~ p))].
\Proof. mv. 
intro p. intro H. box_i. intro H2. apply H. dia_i w0. exact H2. 
\Qed.
\end{Verbatim}

\section{Modal Logics beyond K}

@@

todo: automation, firstorder
